---
title: "Research Project"
author: "Xiao Xiao"
date: "2023-06-27"
output: html_document
---
1. introduction of project

2. run competing models and present in a table

3. avoid screen shots

4. screenreg

Imported data from Kaggle.
This research project is targeted to predict the probability of company bankruptcy based on original 6819 observations with 96 financial indicators of enterprises across Taiwan.
```{r}
df <- read.csv("/Users/xiaoxiao/Downloads/ME315/data.csv")
dim(df)
any(is.na(df))
```

Then, we select and observe some important features according in financial accounting.
```{r}
library(magrittr)
library(dplyr) 

subdf <- df %>% select( Bankrupt., Net.profit.before.tax.Paid.in.capital, Operating.Gross.Margin , Debt.ratio.., Quick.Ratio,Interest.Coverage.Ratio..Interest.expense.to.EBIT. , Inventory.Turnover.Rate..times. ,Total.Asset.Turnover)
summary(subdf)

# separate bankruptcy and non-bankruptcy companies
subdf0 <- subdf %>% filter(Bankrupt.==0)
subdf1 <- subdf %>% filter(Bankrupt.==1)
dim(subdf0)
dim(subdf1)
```

EDA
```{r}
par(mfrow=c(3,2))
hist(subdf$Bankrupt.,breaks=2,xlab = 'Bankruptcy',main = paste("Histogram of Bankruptcy"))
hist(subdf$Inventory.Turnover.Rate..times.,breaks=10,main = paste("Histogram of Total Asset Turnover"))
hist(subdf0$Inventory.Turnover.Rate..times.,breaks=10,main = paste("Histogram of Total Asset Turnover for Non-Bankrupt"))
hist(subdf1$Inventory.Turnover.Rate..times.,breaks=10,main = paste("Histogram of Total Asset Turnover for Bankrupt"))
# not same as Operating Expense Rate
hist(subdf0$Operating.Gross.Margin,breaks=10,main = paste("Histogram of Operating Gross Rate for Non-bankrupt"))
hist(subdf1$Operating.Gross.Margin,breaks=10,main = paste("Histogram of Operating Gross Rate for Bankrupt"))


hist(subdf1$Quick.Ratio,breaks=10,main = paste("Histogram of Operating Gross Rate for Bankrupt"))
hist(subdf0$Quick.Ratio,breaks=10,main = paste("Histogram of Operating Gross Rate for Bankrupt"))


par(mfrow=c(3,2))
plot(density(subdf$Total.Asset.Turnover)) 
plot(density(log(subdf$Total.Asset.Turnover)))#log transformation
plot(density(subdf$Net.profit.before.tax.Paid.in.capital)) 
plot(density(log(subdf$Net.profit.before.tax.Paid.in.capital)))#log transformation
plot(density(subdf$Debt.ratio..)) 
plot(density(log(subdf$Debt.ratio..)))#log transformation


par(mfrow=c(1,2))
boxplot(Operating.Gross.Margin ~ Bankrupt., data = subdf)
boxplot(Net.profit.before.tax.Paid.in.capital~Bankrupt., data = subdf)


plot(df$Bankrupt.,df$Quick.Ratio)
plot(df$Equity.to.Liability,df$Bankrupt.)


pairs(~Bankrupt.+ Net.profit.before.tax.Paid.in.capital + Operating.Gross.Margin + Debt.ratio..+ Quick.Ratio + Interest.Coverage.Ratio..Interest.expense.to.EBIT. + Inventory.Turnover.Rate..times. + Total.Asset.Turnover,df)


df <- df %>% mutate( Log.Total.Asset.Turnover= log(Total.Asset.Turnover), Log.PBT=log(Net.profit.before.tax.Paid.in.capital))
subdf <- subdf %>% mutate(Log.Total.Asset.Turnover= log(Total.Asset.Turnover), Log.PBT=log(Net.profit.before.tax.Paid.in.capital))

```

Unsupervised Learning - PCA
```{r}

subdf <- subdf %>% mutate(Total.Asset.Growth.Rate = df$Total.Asset.Growth.Rate,Cash.Reinvestment..=df$Cash.Reinvestment.., Research.and.development.expense.rate = df$Research.and.development.expense.rate )
```

```{r}
library(ISLR)
apply(subdf[,c(2:11)],MARGIN=2,mean)
sum(is.na(subdf))

pr.out <- prcomp(subdf[,c(2:11)], scale=TRUE, center=T, cor =TRUE)
?prcomp
  #center=T by default
summary(pr.out)
# pr.out$loadings
names(pr.out)
round(pr.out$rotation,3)
head(pr.out$x)


screeplot(pr.out, type="l", main="Screeplot") #kink at 2 components suggests we keep 2
abline(1,0,col='red',lty=2)


pr.out$sdev   #sdev is the standard deviations of the principal components
pr.var <- pr.out$sdev^2
pr.var
# pve <- pr.var/sum(pr.var)
pve <- round(pr.var/sum(pr.var)*100,1)
pve
pdf("pdnf.pdf")
par(mfrow=c(1,2))
plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained",type='b')
plot(cumsum(pve), xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained",type='b')
abline(80,0,col='red',lty=2)

biplot(pr.out, scale=0)
```

Unsupervised Learning - Factor Analysis
```{r}
library(psych)
fit <- factanal(subdf, 3)
names(fit)
print(fit$loadings, digits=2, cutoff=.3, sort=TRUE)

#scree plot
ev <- eigen(cor(subdf)) # get eigenvalues
plot(ev$values,type="l")
```

Supervised Learning - Classification

```{r}
library(ISLR)
df <- df[,c(1:94)]
head(df)
# split the data
set.seed(1)
train=sample(1:nrow(df), 2000)
test = -train
df.train=df[train,]
df.test=df[-train,]
bankrupt.test = df[-train,]$Bankrupt.

library(MASS)

#1. learn model
lda.fit <-  lda(Bankrupt. ~Net.profit.before.tax.Paid.in.capital+ Operating.Gross.Margin + Debt.ratio..+ Quick.Ratio+Interest.Coverage.Ratio..Interest.expense.to.EBIT. + Inventory.Turnover.Rate..times. +Total.Asset.Turnover + Total.Asset.Growth.Rate + Cash.Reinvestment..+ Research.and.development.expense.rate,
              data = df.train)
?lda
#2. predictions on test set
lda.pred <-  predict(lda.fit, df.test)
str(lda.pred)
# str(lda.pred)
print(cbind(lda.pred$class,bankrupt.test))
?cbind
print((lda.pred$class != bankrupt.test))
#3. compute estimate of test error
# MSE is not suitable for Classification
# should use Misclassification Error Rate - proportion of incorrect predictions
lda.fit
mean(lda.pred$class != bankrupt.test)
table(lda.pred$class,bankrupt.test)
```

```{r}
# QDA
qda.fit <-  qda(Bankrupt. ~Net.profit.before.tax.Paid.in.capital+ Operating.Gross.Margin + Debt.ratio..+ Quick.Ratio+Interest.Coverage.Ratio..Interest.expense.to.EBIT. + Inventory.Turnover.Rate..times. +Total.Asset.Turnover+Total.Asset.Growth.Rate + Cash.Reinvestment..+ Research.and.development.expense.rate,
              data = df, subset = train)
round(qda.fit$means,3)
qda.pred <-  predict(qda.fit, df.test)
mean(qda.pred$class != bankrupt.test)

table(qda.pred$class,bankrupt.test)
```
```{r}
# Logistic regression
glm.fit <-  glm(Bankrupt. ~Net.profit.before.tax.Paid.in.capital+ Operating.Gross.Margin + Debt.ratio..+ Quick.Ratio+Interest.Coverage.Ratio..Interest.expense.to.EBIT. + Inventory.Turnover.Rate..times. +Total.Asset.Turnover +Total.Asset.Growth.Rate + Cash.Reinvestment..+ Research.and.development.expense.rate,
              data = df,
              family = binomial,
              subset = train)
summary(glm.fit)
glm.probs <-  predict(glm.fit, df.test, type = "response")
glm.pred <-  rep(0, length(glm.probs))
glm.pred[glm.probs > 0.5] <- 1 # pick a threshold. Here is 0.5
mean(glm.pred != bankrupt.test)

table(glm.pred, bankrupt.test)
```

```{r}
library(class) # to do K-NN
train.X <-  cbind(df$Net.profit.before.tax.Paid.in.capital, df$Operating.Gross.Margin, df$Debt.ratio.., df$Quick.Ratio,df$Interest.Coverage.Ratio..Interest.expense.to.EBIT., df$Inventory.Turnover.Rate..times.,df$Total.Asset.Turnover)[train,]
test.X <-  cbind(df$Net.profit.before.tax.Paid.in.capital, df$Operating.Gross.Margin, df$Debt.ratio.., df$Quick.Ratio,df$Interest.Coverage.Ratio..Interest.expense.to.EBIT., df$Inventory.Turnover.Rate..times.,df$Total.Asset.Turnover)[test,]
train.bankuptcy <-  df$Bankrupt.[train]
set.seed(1)

# KNN (k=1)
knn.pred <-  knn(train.X, test.X, train.bankuptcy, k = 1)
mean(knn.pred != bankrupt.test)

# KNN (k=10)
knn.pred <-  knn(train.X, test.X, train.bankuptcy, k = 10)
mean(knn.pred != bankrupt.test)

# KNN (k=100)
knn.pred <-  knn(train.X, test.X, train.bankuptcy, k = 100)
mean(knn.pred != bankrupt.test)
table(knn.pred,bankrupt.test)
```

